---
title: "Trial and robustness run reference"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Cross-reference for CMP Trial and Robustness Runs</div>

This page provides a quick reference to the procedure settings recorded in
`tuns_ctrl.csv` and `robs_ctrl.csv`. The tables are rebuilt from those files
when the website is rendered, so they remain aligned with the analysis control
files.

## Tuning trials

These are the procedure designs examined during tuning. MP29 and MP32 were
subsequently carried forward for robustness testing.

```{r}
#| label: tbl-tuning-controls
#| tbl-cap: "Settings for the 27 tuning trials. A dash means that the setting does not apply to that rule."
#| echo: false

tuning <- read.csv("../tuns_ctrl.csv", check.names = FALSE,
                   na.strings = c("", "NA"))

show_number <- function(x) {
  out <- format(x, trim = TRUE, scientific = FALSE, digits = 7)
  out[is.na(x)] <- "—"
  out
}

tuning_table <- data.frame(
  MP = tuning$label,
  Rule = sub("^hockeystick$", "Hockeystick", tuning$hcr),
  `Information source` = tuning$estimator,
  `Summary measure` = tuning$metric,
  `Target catch` = show_number(tuning$target),
  `Minimum catch` = show_number(tuning$min),
  Limit = show_number(tuning$lim),
  Trigger = show_number(tuning$trigger),
  `Annual decrease limit` = show_number(tuning$dlow),
  `Annual increase limit` = show_number(tuning$dupp),
  check.names = FALSE
)

tuning_table$Rule <- tools::toTitleCase(tuning_table$Rule)
knitr::kable(tuning_table, align = c("l", "l", "l", "l", rep("r", 6)))
```

## Robustness runs

The robustness analysis applies MP29 and MP32 to nine alternative operating
models. Each row below identifies one procedure–operating-model combination.

```{r}
#| label: tbl-robustness-controls
#| tbl-cap: "MP29 and MP32 robustness runs by operating model."
#| echo: false

robustness <- read.csv("../robs_ctrl.csv", check.names = FALSE,
                       na.strings = c("", "NA"))

robustness_table <- data.frame(
  `Run name` = robustness$run,
  MP = robustness$label,
  `Operating model` = robustness$om,
  Rule = tools::toTitleCase(sub("\\.hcr$", "", robustness$hcr)),
  `Information source` = robustness$estimator,
  `Target catch` = show_number(robustness$target),
  `Minimum catch` = show_number(robustness$min),
  Limit = show_number(robustness$lim),
  Trigger = show_number(robustness$trigger),
  check.names = FALSE
)

knitr::kable(robustness_table,
             align = c("l", "l", "l", "l", "l", rep("r", 4)))
```

## Reading the tables

- **Rule** is the form of the harvest control rule.
- **Information source** identifies the stock information supplied to the
  rule. `all.idx` means that all specified abundance indices are combined.
- **Limit** is the indicator value at or below which the minimum catch applies.
- **Trigger** is the indicator value at or above which the target catch applies.
- **Annual decrease and increase limits** are multipliers on the previous
  year's catch advice. For example, 0.85 and 1.20 limit a change to a 15%
  decrease or a 20% increase.
- **Operating model** identifies the alternative description of stock dynamics
  used for robustness testing. See the MSE analysis for the assumptions that
  define each operating model.

The detailed definitions of the two advanced candidates are on the
[MP29](mp29.qmd) and [MP32](mp32.qmd) pages.
