---
title: "Earlier trial and robustness controls"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Earlier CMP Trial and Robustness Controls</div>

::: {.status-note}
**Historical development record.** The controls on this page include earlier
100-draw settings and candidates that are not in the named SC14 release. They
must not be used as the current CMP list. Use the [SC14 CMP
registry](index.qmd) and [release agreement report](../governance/release-check.qmd)
for the current eight-CMP set.
:::

This page preserves the procedure settings recorded in the earlier
`tuns_ctrl.csv` and `robs_ctrl.csv` files so that the development sequence can
still be understood.

## Tuning trials

These are procedure designs examined during earlier tuning. The table is not a
current release control file.

```{r}
#| label: tbl-tuning-controls
#| tbl-cap: "Settings for the 29 tuning trials. A dash means that the setting does not apply to that rule."
#| echo: false

tuning <- read.csv("../tuns_ctrl.csv", check.names = FALSE,
                   na.strings = c("", "NA"))

show_number <- function(x) {
  out <- format(x, trim = TRUE, scientific = FALSE, digits = 7)
  out[is.na(x)] <- "—"
  out
}

tuning_table <- data.frame(
  MP = tuning$label,
  Rule = sub("^hockeystick$", "Hockeystick", tuning$hcr),
  `Information source` = tuning$estimator,
  `Summary measure` = tuning$metric,
  `Target catch` = show_number(tuning$target),
  `Minimum catch` = show_number(tuning$min),
  Limit = show_number(tuning$lim),
  Trigger = show_number(tuning$trigger),
  `Annual decrease limit` = show_number(tuning$dlow),
  `Annual increase limit` = show_number(tuning$dupp),
  check.names = FALSE
)

tuning_table$Rule <- tools::toTitleCase(tuning_table$Rule)
knitr::kable(tuning_table, align = c("l", "l", "l", "l", rep("r", 6)))
```

## Robustness runs

The earlier robustness table applied MP29, MP32, MP43, and MP44 to nine
alternative operating models. Each row below identifies one historical
procedure–operating-model combination. The current 500-draw release instead
contains all eight SC14 CMPs, listed in the release run index.

```{r}
#| label: tbl-robustness-controls
#| tbl-cap: "MP29, MP32, MP43, and MP44 robustness runs by operating model."
#| echo: false

robustness <- read.csv("../robs_ctrl.csv", check.names = FALSE,
                       na.strings = c("", "NA"))

robustness_table <- data.frame(
  `Run name` = robustness$run,
  MP = robustness$label,
  `Operating model` = robustness$om,
  Rule = tools::toTitleCase(sub("\\.hcr$", "", robustness$hcr)),
  `Information source` = robustness$estimator,
  `Target catch` = show_number(robustness$target),
  `Minimum catch` = show_number(robustness$min),
  Limit = show_number(robustness$lim),
  Trigger = show_number(robustness$trigger),
  check.names = FALSE
)

knitr::kable(robustness_table,
             align = c("l", "l", "l", "l", "l", rep("r", 4)))
```

## Reading the tables

- **Rule** is the form of the harvest control rule.
- **Information source** identifies the stock information supplied to the
  rule. `all.idx` means that all specified abundance indices are combined.
- **Limit** is the indicator value at or below which the minimum catch applies.
- **Trigger** is the indicator value at or above which the target catch applies.
- **Annual decrease and increase limits** are multipliers on the previous
  year's catch advice. For example, 0.85 and 1.20 limit a change to a 15%
  decrease or a 20% increase.
- **Operating model** identifies the alternative description of stock dynamics
  used for robustness testing. See the MSE analysis for the assumptions that
  define each operating model.

The [MP29](mp29.qmd) and [MP32](mp32.qmd) pages document their current 500-draw
sensitivity specifications. The registry and SC14 report record the complete
eight-CMP refinement.
