---
title: "Checks of agreement among the SC14 files"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Checks of Agreement among the SC14 Files</div>

```{r}
#| label: release-check-setup
#| include: false

manifest <- yaml::read_yaml("../../release/sc14-mse-2026-rc1.yml")
checks <- read.csv("../../release/sc14-check-results.csv",
  stringsAsFactors = FALSE, check.names = FALSE)
run_index <- read.csv("../../release/sc14-run-index.csv",
  stringsAsFactors = FALSE, check.names = FALSE)
file_list <- read.csv("../../release/sc14-file-list.csv",
  stringsAsFactors = FALSE, check.names = FALSE)
```

## Result

This page reports automatic checks for **`r manifest$release_id`**. It answers
the reviewer's question of whether the CMP settings, named simulation runs,
results, figures, downloads, Slick file, and papers belong to the same
release.

```{r}
#| label: tbl-release-check-summary
#| tbl-cap: "Summary of the automatic checks."
#| echo: false

summary <- as.data.frame(table(factor(checks$outcome,
  levels = c("Pass", "Not ready", "Fail"))))
names(summary) <- c("Outcome", "Number of checks")
knitr::kable(summary)
```

**Pass** means that the automatic comparison found agreement. **Not ready**
means that an administrative release step, such as committing the named
revision or rendering a file, remains. **Fail** means that scientific or
numerical content differs and must be resolved.

## Checks requiring attention

```{r}
#| label: tbl-release-check-attention
#| tbl-cap: "Checks that are not yet ready or failed."
#| echo: false

attention <- checks[checks$outcome != "Pass",
  c("area", "check", "outcome", "observed", "note")]
if (!nrow(attention)) {
  attention <- data.frame(Result = "All automatic checks passed.")
}
knitr::kable(attention)
```

## Complete check record

```{r}
#| label: tbl-release-check-complete
#| tbl-cap: "Complete automatic check record."
#| echo: false

knitr::kable(checks[c("area", "check", "outcome", "expected", "observed")])
```

## Simulation runs included

The release contains `r nrow(run_index)` CMP, operating-model, and stock
combinations. Two-stock operating models have separate rows for the northern
and southern components.

```{r}
#| label: tbl-release-run-summary
#| tbl-cap: "Summary of the named simulation runs."
#| echo: false

run_summary <- aggregate(run ~ set + om_code + stock + draws,
  run_index, function(x) length(unique(x)))
names(run_summary)[names(run_summary) == "run"] <- "CMP runs"
knitr::kable(run_summary)
```

[Download the complete run list](../data/release/sc14-run-index.csv), the
[complete file list with SHA-256 fingerprints](../data/release/sc14-file-list.csv),
and the [machine-readable check results](../data/release/sc14-check-results.json).

## What the automatic report does not decide

These checks establish that the named files agree. They do not decide whether
a CMP should be shortlisted or adopted, whether an operating model is
plausible, or which management trade-offs should be preferred. Those remain
questions for SC14 and, later, the Commission.

Allocation is recorded only to confirm that the same implemented method is
described throughout the package. This report does not reopen allocation as an
analytical comparison.

Manual review is still required for the scientific interpretation, figure
readability, links, HTML accessibility, and PDF accessibility.
