---
title: "Load candidates into Slick"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Explore MP29 and MP32 Results in Slick</div>

Slick is an interactive tool for exploring MSE results. The simulation is run
in `jmMSE`; this repository converts the resulting performance table into a
Slick file.

## What is needed

The conversion can use either:

- the table of performance results produced by `jmMSE`; or
- a saved `jmMSE` results object that already contains those performance results.

The results must include MP29 and MP32, the operating-model and stock names,
simulation year and replicate, and the calculated performance measures. At a
minimum, the results must include spawning biomass relative to $SB_{MSY}$,
fishing mortality relative to $F_{MSY}$, and catch. Interannual catch change
(IACC) is included when available.

The current reference comparison has been completed for MP29 and MP32 with 100
simulation replicates through 2050. It is stored locally in `jmMSE` as:

- `model/candidates/reference/runs.rds`: the complete simulation object; and
- `model/candidates/reference/performance.rds`: the smaller table used for
  Slick.

These are reference-operating-model results. They are suitable for checking the
candidate calculations and comparing MP29 with MP32, but they are not a
substitute for robustness testing across the full set of operating models.

The nine-OM robustness set has also been completed and is stored locally in
`jmMSE/model/candidates/robustness/`. It contains MP29 and MP32, 100 simulation
replicates, and annual results through 2050 for every operating model.

## Run or extend the reference comparison

From the `jmMSE` folder on branch `devel`:

```sh
Rscript run_candidate_mps.R \
  --candidates=tun29,tun32 \
  --mode=reference \
  --cores=1
```

The candidate codes come from `candidate_mps.csv`. Additional hockeystick or
powerramp trials can be included for contrast, for example:

```sh
Rscript run_candidate_mps.R \
  --candidates=tun25,tun29,tun32 \
  --mode=reference \
  --cores=1
```

The runner uses the estimator and rule implementations saved with the original
tuning runs, then applies the selected rows from `candidate_mps.csv`. This
preserves the calculations used to tune the candidates. Moving to a newer
combined-index implementation should be treated as a method change requiring
the operating models to be rebuilt and the candidates to be retuned and
revalidated.

The validated reference run used R 4.6.1 with FLCore 2.6.30, FLFishery 0.4.0,
mse 2.4.8.9004, msemodules 0.2.2, and FLjjm 0.3.4. The run folder contains the
complete `sessionInfo.txt` record.

## Prepare the Slick file

From the main `jmMSE26` folder:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

slick <- build_candidate_slick(
  performance_file = "../jmMSE/model/candidates/reference/performance.rds",
  out_file = "output/jm_candidates.slick"
)
```

The script organizes the results by:

- management procedure (MP29 or MP32);
- operating model and stock;
- simulation replicate and year; and
- performance measure, including Kobe-plot values for $SB/SB_{MSY}$ and
  $F/F_{MSY}$.

Before saving the Slick file, the script checks that the expected dimensions
and labels agree. The technical checks are shown in the code for analysts who
maintain the export.

The current file contains 100 simulations, one reference operating model, two
management procedures, four performance measures, and 26 annual values from
2025 through 2050.

## Prepare the robustness Slick file

The robustness operating models and candidate runs can be reproduced from
`jmMSE` branch `devel` with:

```sh
Rscript prepare_candidate_robustness_oms.R
Rscript run_candidate_mps.R \
  --candidates=tun29,tun32 \
  --mode=robustness \
  --cores=1
```

Then build the Slick file from `jmMSE26`:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

robustness_slick <- build_candidate_slick(
  performance_file = "../jmMSE/model/candidates/robustness/performance.rds",
  out_file = "output/jm_candidates_robustness.slick"
)
```

The Slick file contains the nine operating models as 13 OM–stock series. Five
operating models contain one stock. Four contain separate Southern and Northern
stocks, and Slick retains both stock results rather than averaging them. The
dimensions are 100 simulations, 13 OM–stock series, two management procedures,
four performance measures, and 26 annual values from 2025 through 2050.

Do not combine results produced under different package sets without first
confirming that the candidate calculations are equivalent.

## Inspect or launch

```{r}
#| eval: false
library(Slick)

slick <- readRDS("output/jm_candidates.slick")
Check(slick)
methods::validObject(slick)

dim(Value(Timeseries(slick)))
Metadata(MPs(slick))
Design(slick)

App(slick = slick)
```

## What to save with the Slick file

Save the Slick file together with the original performance results, their
digital fingerprint, the exact saved versions of `jmMSE` and `jmMSE26`, and the
versions of the R packages used.
