---
title: "Load candidates into Slick"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Load the MP29/MP32 Candidate Set into Slick</div>

The project-local adapter `R/build_candidate_slick.R` converts current `jmMSE`
performance outputs into a validated Slick object. It does not run the MSE and
does not fabricate missing results.

## Required input

Use either:

- a long performance `data.frame`/`data.table`; or
- an `FLmses`-like S4 object with a populated `performance` slot.

The input must contain MP29 and MP32 runs and the columns `run`, `om`, `biol`,
`statistic`, `year`, `iter`, and `data`. At minimum, statistics `SBMSY`, `FMSY`,
and `C` must be present. `IACC` is included when available.

The current checkout does not yet contain MP29/MP32 result artifacts in
`model/tune/` or `model/robustness/`; the available top-level tuning artifacts
end at MP22. Run the current candidate/robustness workflow before attempting
the production export.

## Build the Slick object

From the `jmMSE26` repository root:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

slick <- build_candidate_slick(
  performance_file = "../jmMSE/model/robustness/robs.rds",
  out_file = "output/jm_candidates.slick"
)
```

The builder creates:

- `MPs()` metadata for MP29 and MP32 from the registry;
- `OMs()` metadata for each OM/stock combination;
- `Timeseries()` values ordered as iteration x OM x MP x indicator x year; and
- `Kobe()` values for `SB/SBMSY` and `F/FMSY` over projection years.

It calls `Check()` on the time-series, Kobe, and final Slick objects, calls
`methods::validObject()`, and writes the `.slick` file only after validation.

## Inspect or launch

```{r}
#| eval: false
library(Slick)

slick <- readRDS("output/jm_candidates.slick")
Check(slick)
methods::validObject(slick)

dim(Value(Timeseries(slick)))
Metadata(MPs(slick))
Design(slick)

App(slick = slick)
```

## Reproducibility record

Archive the `.slick` file together with the source performance artifact, its
checksum, the `jmMSE` commit, the `jmMSE26` commit, and the R package versions.

