---
title: "Load candidates into Slick"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Explore the Eight CMP Comparison Cases in Slick</div>

Slick is an interactive tool for exploring MSE results. The simulation is run
in `jmMSE`; this repository converts the resulting performance table into a
Slick file.

## Open the current file

### In the hosted app

1. [Download `jm_candidates.slick`](https://raw.githubusercontent.com/SPRFMO/jmMSE26/main/output/jm_candidates.slick).
2. Open the [Blue Matter Slick app](https://shiny.bluematterscience.com/app/slick).
3. If the app has stopped, select **Restart app**.
4. Under **Load your MSE Results**, select the downloaded `.slick` file.
5. When the Overview appears, use the MP and OM filters to explore the results.

The hosted app accepts file uploads but does not currently provide a field for
loading a file directly from a GitHub URL.

### From R

```{r}
#| eval: false
install.packages("Slick") # needed only once
library(Slick)

url <- paste0(
  "https://raw.githubusercontent.com/SPRFMO/jmMSE26/",
  "main/output/jm_candidates.slick"
)
file <- tempfile(fileext = ".slick")
download.file(url, file, mode = "wb")

slick <- readRDS(file)
Check(slick)
App(slick = slick)
```

### Make it a selectable case study

Slick's case-study menu is populated from the
[`Blue-Matter/SlickLibrary`](https://github.com/Blue-Matter/SlickLibrary)
repository. To request direct integration, copy the file to
`Slick_Objects/SPRFMO_Jack_Mackerel.slick` in a fork of that repository and
submit a pull request. Once the SlickLibrary maintainers merge it, **SPRFMO
Jack Mackerel** will appear in the hosted app's case-study menu. See the
[official contribution instructions](https://slick.bluematterscience.com/articles/DevelopersGuide.html#slick-object-library).

## What is needed

The conversion can use either:

- the table of performance results produced by `jmMSE`; or
- a saved `jmMSE` results object that already contains those performance results.

The results must include the eight CMPs listed below, the operating-model and stock names,
simulation year and replicate, and the calculated performance measures. At a
minimum, the results must include spawning biomass relative to $SB_{MSY}$,
fishing mortality relative to $F_{MSY}$, and catch. Interannual catch change
(IACC) is included when available.

The current reference comparison has been completed for eight CMPs with 100
simulation replicates through 2050. It is stored locally in `jmMSE` as:

- `model/candidates/reference/runs.rds`: the complete simulation object; and
- `model/candidates/reference/performance.rds`: the smaller table used for
  Slick.

These are reference-operating-model results. They are suitable for checking the
candidate calculations and comparing the rule and data contrasts, but they are not a
substitute for robustness testing across the full set of operating models.

The nine-OM robustness set has also been completed and is stored locally in
`jmMSE/model/candidates/robustness/`. It contains all eight CMPs, 100 simulation
replicates, and annual results through 2050 for every operating model.

## CMPs included

MP29 and MP32 remain the two advanced candidates. The other six procedures are
comparison cases that help show which part of a procedure causes a change in
performance.

| CMP | Role in the comparison |
|---|---|
| MP29 | Optimal hockeystick case: target 2000, limit 0.1, all selected indices. |
| MP32 | Alternative powerramp case: target 1500, limit 0.1, all selected indices. |
| MP18 | Extreme high-target case: hockeystick target 5000. |
| MP24 | Low-target case: hockeystick target 1750 and limit 0. |
| MP31 | Higher-limit case: hockeystick target 2000 and limit 0.3. |
| MP23 | Different rule shape: slope HCR with its original five-year averaging window. |
| MP35 | CPUE-only powerramp contrast: target 1500 and limit 0.1. |
| MP36 | CPUE-only hockeystick contrast: target 2000 and limit 0.1. |

The MP filter includes one-click selections for the advanced candidates,
target contrasts, the limit contrast, rule-shape contrasts, CPUE-only
contrasts, all CMPs, and each individual CMP.

MP23 retains the five-year averaging window used in its original tuning code.
MP29 and MP32 use three years, so the MP23 comparison changes both rule shape
and averaging window; it should not be interpreted as a pure test of shape
alone.

## Run or extend the reference comparison

From the `jmMSE` folder on branch `devel`:

```sh
Rscript run_candidate_mps.R \
  --candidates=tun29,tun32,tun18,tun24,tun31,tun23,tun35,tun36 \
  --mode=reference \
  --cores=1
```

The candidate codes come from `candidate_mps.csv`. Additional hockeystick or
powerramp trials can be included for contrast, for example:

```sh
Rscript run_candidate_mps.R \
  --candidates=tun25,tun29,tun32 \
  --mode=reference \
  --cores=1
```

The runner uses the recorded definitions in `candidate_mps.csv` together with
the original estimator and rule functions in `jmMSE`. Shortcut-assessment
errors are generated from the original error model using fixed seeds so the
comparison can be repeated. Each CMP starts from the same fixed random stream
within an operating model, so adding or reordering CMPs does not change the
other CMP results. Moving to a newer
combined-index implementation should be treated as a method change requiring
the operating models to be rebuilt and the candidates to be retuned and
revalidated.

The validated reference run used R 4.6.1 with FLCore 2.6.30, FLFishery 0.4.0,
mse 2.4.8.9004, msemodules 0.2.2, and FLjjm 0.3.4. The run folder contains the
complete `sessionInfo.txt` record.

## Prepare the combined Slick file

From the main `jmMSE26` folder:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

reference_slick <- build_candidate_slick(
  performance_file = "../jmMSE/model/candidates/reference/performance.rds",
  out_file = NULL,
  historical_om_files = candidate_reference_om_files("../jmMSE")
)
```

The script organizes the results by:

- management procedure (one of the eight CMP comparison cases);
- operating model and stock;
- simulation replicate and year; and
- performance measure, including Kobe-plot values for $SB/SB_{MSY}$ and
  $F/F_{MSY}$.

It also prepares three summary views using the comparison periods already used
in `jmMSE`:

- **Boxplots** show the distribution among simulation replicates in the final
  year, 2050.
- **Quilts** compare stock status, fishing pressure, catch, and catch stability
  among operating models and MPs. Long-term results use 2036–2040; short-term
  catch uses 2026–2030.
- **Trade-off plots** use the same periods to compare conservation, catch, and
  catch-stability measures directly.

Kobe quadrant probabilities are calculated from $SB/SB_{MSY}$ and $F/F_{MSY}$.
Non-finite ratios, which can occur when a stock component approaches zero, are
treated as missing rather than replaced with invented values.

Before saving the Slick file, the script checks that the expected dimensions
and labels agree. The technical checks are shown in the code for analysts who
maintain the export.

The reference is labelled `om11` in Slick. Its source name, `h1_0.16`, remains
in the OM metadata for traceability.

### Add the robustness sets

The robustness operating models and candidate runs can be reproduced from
`jmMSE` branch `devel` with:

```sh
Rscript prepare_candidate_robustness_oms.R
Rscript run_candidate_mps.R \
  --candidates=tun29,tun32,tun18,tun24,tun31,tun23,tun35,tun36 \
  --mode=robustness \
  --cores=1
```

Then build the robustness component and combine it with the reference:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

robustness_slick <- build_candidate_slick(
  performance_file = "../jmMSE/model/candidates/robustness/performance.rds",
  out_file = NULL,
  historical_om_files = candidate_robustness_om_files("../jmMSE")
)

slick <- build_combined_candidate_slick(
  reference_slick = reference_slick,
  robustness_slick = robustness_slick,
  out_file = "output/jm_candidates.slick"
)
```

The resulting Slick file contains 14 OM–stock series: one reference, five
single-stock robustness cases, and eight stock-specific series from four
two-stock robustness cases. Slick retains the Southern and Northern results
rather than averaging them.

The `Set` filter provides three main categories:

- **Reference**: om11;
- **Robustness CJM**: the five single-stock robustness OMs; and
- **Robustness 2-stock**: the Southern and Northern results from the four
  two-stock robustness OMs.

The OM and Stock filters remain available for selecting individual models and
stock components. One-click presets provide quick selections for the reference,
all CJM cases, the CJM robustness cases, the two-stock robustness cases, all
OMs, and each individual alternative OM. Selecting a two-stock OM includes both
its Southern and Northern rows.

The dimensions are 100 simulations, 14 OM–stock series, eight management
procedures, four performance measures, and 81 annual values from 1970 through
2050. Each series uses its own conditioned history through 2025; projections
under the eight CMPs begin in 2026.

Do not combine results produced under different package sets without first
confirming that the candidate calculations are equivalent.

## Inspect or launch

```{r}
#| eval: false
library(Slick)

slick <- readRDS("output/jm_candidates.slick")
Check(slick)
methods::validObject(slick)

dim(Value(Timeseries(slick)))
Metadata(MPs(slick))
Design(slick)

App(slick = slick)
```

## What to save with the Slick file

Save the Slick file together with the original performance results, their
digital fingerprint, the exact saved versions of `jmMSE` and `jmMSE26`, and the
versions of the R packages used.
