---
title: "Load candidates into Slick"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Explore MP29 and MP32 Results in Slick</div>

Slick is an interactive tool for exploring MSE results. The supplied conversion
script prepares MP29 and MP32 results for Slick after the MSE runs have been
completed. It does not run the MSE and it does not fill in missing results.

## What is needed

The conversion can use either:

- the table of performance results produced by `jmMSE`; or
- a saved `jmMSE` results object that already contains those performance results.

The results must include MP29 and MP32, the operating-model and stock names,
simulation year and replicate, and the calculated performance measures. At a
minimum, the results must include spawning biomass relative to $SB_{MSY}$,
fishing mortality relative to $F_{MSY}$, and catch. Interannual catch change
(IACC) is included when available.

The current working folder does not yet contain saved MP29/MP32 results in
`model/tune/` or `model/robustness/`; the available top-level tuning artifacts
end at MP22. Run the current candidate and robustness analyses before attempting
the final Slick export.

## Prepare the Slick file

From the main `jmMSE26` folder:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

slick <- build_candidate_slick(
  performance_file = "../jmMSE/model/robustness/robs.rds",
  out_file = "output/jm_candidates.slick"
)
```

The script organizes the results by:

- management procedure (MP29 or MP32);
- operating model and stock;
- simulation replicate and year; and
- performance measure, including Kobe-plot values for $SB/SB_{MSY}$ and
  $F/F_{MSY}$.

Before saving the Slick file, the script checks that the expected dimensions
and labels agree. The technical checks are shown in the code for analysts who
maintain the export.

## Inspect or launch

```{r}
#| eval: false
library(Slick)

slick <- readRDS("output/jm_candidates.slick")
Check(slick)
methods::validObject(slick)

dim(Value(Timeseries(slick)))
Metadata(MPs(slick))
Design(slick)

App(slick = slick)
```

## What to save with the Slick file

Save the Slick file together with the original performance results, their
digital fingerprint, the exact saved versions of `jmMSE` and `jmMSE26`, and the
versions of the R packages used.
