---
title: "Load candidates into Slick"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">Explore the Eight CMP Comparison Cases in Slick</div>

Slick is an interactive tool for exploring MSE results. The simulation is run
in `jmMSE`; this repository converts the resulting performance table into a
Slick file.

## Open the current file

### In the hosted app

1. [Download the current `jm_candidates.slick`](https://raw.githubusercontent.com/SPRFMO/jmMSE26/main/output/jm_candidates.slick).
2. Open the [Blue Matter Slick app](https://shiny.bluematterscience.com/app/slick).
3. If the app has stopped, select **Restart app**.
4. Under **Load your MSE Results**, select the downloaded `.slick` file.
5. When the Overview appears, use the MP and OM filters to explore the results.
   The current file opens with **All CMPs** selected, so all eight annual-limit
   variants should
   be visible immediately.

If only HS+20 and PR+20 appear, open the **MPs** filter and select **All CMPs**.
That indicates either that the **Central rule-shape comparison** preset is active or
that an older copy of the same-named file was loaded. The current file contains
eight CMPs. Its SHA-256 fingerprint is recorded in the traceability page for
the published branch.

The hosted app accepts file uploads but does not currently provide a field for
loading a file directly from a GitHub URL.

### From R

```{r}
#| eval: false
install.packages("Slick") # needed only once
library(Slick)

url <- paste0(
  "https://raw.githubusercontent.com/SPRFMO/jmMSE26/",
  "main/",
  "output/jm_candidates.slick"
)
file <- tempfile(fileext = ".slick")
download.file(url, file, mode = "wb")

slick <- readRDS(file)
Check(slick)
App(slick = slick)
```

### Make it a selectable case study

Slick's case-study menu is populated from the
[`Blue-Matter/SlickLibrary`](https://github.com/Blue-Matter/SlickLibrary)
repository. To request direct integration, copy the file to
`Slick_Objects/SPRFMO_Jack_Mackerel.slick` in a fork of that repository and
submit a pull request. Once the SlickLibrary maintainers merge it, **SPRFMO
Jack Mackerel** will appear in the hosted app's case-study menu. See the
[official contribution instructions](https://slick.bluematterscience.com/articles/DevelopersGuide.html#slick-object-library).

## What is needed

The conversion can use either:

- the table of performance results produced by `jmMSE`; or
- a saved `jmMSE` results object that already contains those performance results.

The results must include the eight CMPs listed below, the operating-model and stock names,
simulation year and replicate, and the calculated performance measures. At a
minimum, the results must include spawning biomass relative to $SB_{MSY}$,
fishing mortality relative to $F_{MSY}$, and catch. Interannual catch change
(IACC) is included when available. The published Slick file also requires
vulnerable biomass relative to its 2025 level and relative to equilibrium
vulnerable biomass at the fishing mortality that produces MSY.

The current published comparison uses the completed 100-posterior-iteration
runs through 2050. The eight CMPs are retained locally in `jmMSE` across:

- `model/candidates/reference/`: HS+20, HS-20, PR+20, and PR-20; and
- `model/candidates/additional-change-limits/reference/`: HSsym, HS-30,
  PRsym, and PR-30.

The augmented six-metric tables used for the published Slick file are retained
in this repository as
`output/candidate-performance/reference/performance_with_vb.rds` and
`output/candidate-performance/robustness/performance_with_vb.rds`.

These are reference-operating-model results. They are suitable for checking the
candidate calculations and comparing the rule and data contrasts, but they are not a
substitute for robustness testing across the full set of operating models.

The corresponding nine-OM robustness results are stored in
`jmMSE/model/candidates/robustness/` and
`jmMSE/model/candidates/additional-change-limits/robustness/`. Together they
contain all eight CMPs, 100 posterior iterations, and annual results through
2050 for every operating model.

## CMPs included

HS+20 (MP29) and PR+20 (MP32) anchor the rule-shape comparison. Six retuned
annual-limit variants test reverse-asymmetric, symmetric, and 30%-decrease
constraints within the same two HCR families.

| CMP | Role in the comparison |
|---|---|
| HS+20 (MP29) | Hockey-stick; -15%/+20% annual limits; trigger 2.125. |
| HS-20 (MP43) | Hockey-stick; -20%/+15% annual limits; trigger 3.0625. |
| HSsym (MP45) | Hockey-stick; -15%/+15% annual limits; trigger 3.0625. |
| HS-30 (MP47) | Hockey-stick; -30%/+20% annual limits; trigger 3.0625. |
| PR+20 (MP32) | Power ramp; -15%/+20% annual limits; trigger 1.421875. |
| PR-20 (MP44) | Power ramp; -20%/+15% annual limits; trigger 1.9140625. |
| PRsym (MP46) | Power ramp; -15%/+15% annual limits; trigger 1.9375. |
| PR-30 (MP48) | Power ramp; -30%/+20% annual limits; trigger 1.890625. |

The MP filter includes one-click selections for the central rule-shape pair,
each HCR family, each annual-limit design, all CMPs, and each individual CMP.
The six annual-limit variants were retuned under reference OM11 toward a mean
dynamic Kobe-green probability of 0.60 over 2041–2050 using 100 posterior
iterations. Recalculation from the archived run objects gives 0.262 for HS+20
and 0.273 for PR+20, so those two initial configurations are not equally tuned
to the six variants in this 100-draw export. A separate completed 500-draw
refinement reaches 0.60 (within 0.01) for all eight CMPs, but should replace
this file only when its corresponding reference and robustness performance
artifacts are adopted together.

## Run or extend the reference comparison

From the `jmMSE` folder on branch `dev_jim`:

```sh
Rscript run_candidate_mps.R \
  --candidates=tun29,tun43,tun45,tun47,tun32,tun44,tun46,tun48 \
  --mode=reference \
  --cores=1
```

The candidate codes come from `candidate_mps.csv`. Additional hockeystick or
powerramp trials can be included for contrast, for example:

```sh
Rscript run_candidate_mps.R \
  --candidates=tun25,tun29,tun32 \
  --mode=reference \
  --cores=1
```

The runner uses the recorded definitions in `candidate_mps.csv` together with
the original estimator and rule functions in `jmMSE`. Shortcut-assessment
errors are generated from the original error model using fixed seeds so the
comparison can be repeated. Each CMP starts from the same fixed random stream
within an operating model, so adding or reordering CMPs does not change the
other CMP results. Moving to a newer
combined-index implementation should be treated as a method change requiring
the operating models to be rebuilt and the candidates to be retuned and
revalidated.

The validated reference run used R 4.6.1 with FLCore 2.6.30, FLFishery 0.4.0,
mse 2.4.8.9004, msemodules 0.2.2, and FLjjm 0.3.4. The run folder contains the
complete `sessionInfo.txt` record.

## Prepare the combined Slick file

From the main `jmMSE26` folder:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

reference_slick <- build_candidate_slick(
  performance_file =
    "output/candidate-performance/reference/performance_with_vb.rds",
  out_file = NULL,
  historical_om_files = candidate_reference_om_files("../jmMSE"),
  require_vb = TRUE
)
```

The script organizes the results by:

- management procedure (one of the eight CMP comparison cases);
- operating model and stock;
- simulation replicate and year; and
- performance measure, including conventional $SB/SB_{MSY}$, the dynamic
  spawning-biomass ratio used for tuning, $F/F_{MSY}$, plus
  $VB/VB_{2025}$ and $VB/VB_{\mathrm{MSY}}$.

It also prepares three summary views using the comparison periods already used
in `jmMSE`:

- **Boxplots** show the distribution among simulation replicates in the final
  year, 2050.
- **Quilts** compare stock status, fishing pressure, catch, and catch stability
  among operating models and MPs. They also include both vulnerable-biomass
  ratios. Long-term results use the SC14 tuning window, 2041–2050;
  short-term catch uses 2026–2030.
- **Trade-off plots** use the same periods to compare conservation, catch, and
  catch-stability measures directly, including both vulnerable-biomass ratios.

The vulnerable-biomass time series are available from 2024 through 2050;
earlier years are intentionally missing. `VB/VB[2025]` uses each simulated
stock's 2025 vulnerable biomass as its denominator. `VB/VB[MSY]` uses
equilibrium vulnerable biomass at the fishing mortality that produces MSY,
not MSY catch.

Kobe quadrant probabilities are calculated from the dynamic spawning-biomass
ratio and $F/F_{MSY}$. The dynamic denominator applies the equilibrium
$SB_{MSY}/SB_0$ fraction to projected unfished spawning biomass in each year
and iteration, matching the SC14 tuning statistic. Conventional
$SB/SB_{MSY}$ remains available as a separate time series and boxplot metric.
When $F_{MSY}$ is zero, $F/F_{MSY}$ is undefined and is treated as missing.
Finite $F/F_{MSY}$ values above 4 are shown at 4 so that a small number of
post-collapse numerical values do not make the plots unreadable. This is a
display rule only: the original `jmMSE` performance tables are not changed.

The OM metadata is kept in exactly the same order as the OM dimension in each
Slick value array. This ensures that Northern and Southern stock results retain
the correct labels in the two-stock operating models.

Before saving the Slick file, the script checks that the expected dimensions
and labels agree. The technical checks are shown in the code for analysts who
maintain the export.

The reference is labelled `om11` in Slick. Its source name, `h1_0.16`, remains
in the OM metadata for traceability.

### Add the robustness sets

The robustness operating models and candidate runs can be reproduced from
the recorded 100-iteration workflow on the `jmMSE` `dev_jim` branch.

```sh
Rscript prepare_candidate_robustness_oms.R
Rscript run_candidate_mps.R \
  --candidates=tun29,tun43,tun45,tun47,tun32,tun44,tun46,tun48 \
  --mode=robustness \
  --cores=1
```

Then build the robustness component and combine it with the reference:

```{r}
#| eval: false
source("R/build_candidate_slick.R")

robustness_slick <- build_candidate_slick(
  performance_file =
    "output/candidate-performance/robustness/performance_with_vb.rds",
  out_file = NULL,
  historical_om_files = candidate_robustness_om_files("../jmMSE"),
  require_vb = TRUE
)

slick <- build_combined_candidate_slick(
  reference_slick = reference_slick,
  robustness_slick = robustness_slick,
  out_file = "output/jm_candidates.slick"
)
```

The resulting Slick file contains 14 OM–stock series: one reference, five
single-stock robustness cases, and eight stock-specific series from four
two-stock robustness cases. Slick retains the Southern and Northern results
rather than averaging them.

The `Set` filter provides three main categories:

- **Reference**: om11;
- **Robustness CJM**: the five single-stock robustness OMs; and
- **Robustness 2-stock**: the Southern and Northern results from the four
  two-stock robustness OMs.

The OM and Stock filters remain available for selecting individual models and
stock components. One-click presets provide quick selections for the reference,
all CJM cases, the CJM robustness cases, the two-stock robustness cases, all
OMs, and each individual alternative OM. Selecting a two-stock OM includes both
its Southern and Northern rows.

The dimensions are 100 posterior iterations, 14 OM–stock series, eight management
procedures, seven performance measures, and 81 annual values from 1970 through
2050. The seven measures are $SB/SB_{MSY}$, dynamic $SB/SB_{MSY}$,
$F/F_{MSY}$, catch, IACC, $VB/VB_{2025}$, and $VB/VB_{\mathrm{MSY}}$. Each series uses its own
conditioned history through 2025; projections under the eight CMPs begin in
2026.

Do not combine results produced under different package sets without first
confirming that the candidate calculations are equivalent.

## Inspect or launch

```{r}
#| eval: false
library(Slick)

slick <- readRDS("output/jm_candidates.slick")
Check(slick)
methods::validObject(slick)

dim(Value(Timeseries(slick)))
Metadata(MPs(slick))
Design(slick)

App(slick = slick)
```

## What to save with the Slick file

Save the Slick file together with the original performance results, their
digital fingerprint, the exact saved versions of `jmMSE` and `jmMSE26`, and the
versions of the R packages used.
