---
title: "Illustrative 2027 TAC calculation"
---

{{< include ../_includes/sprfmo-banner.qmd >}}

<div class="paper-title">How the Candidate Procedures Would Produce 2027 Advice</div>

::: {.status-note}
**This is a calculation plan, not a 2027 TAC recommendation.** None of the
candidate management procedures (CMPs) has been adopted. A numerical
recommendation cannot be produced until the complete annual index data, the
previous catch-advice value, and the rules for data timing, missing values,
rounding, and allocation have been confirmed.
:::

This page shows how the data available for the 2027 decision would be passed
through each CMP. It also identifies where the approximately 630 kt estimate
from the 2026 northern Chile acoustic survey could, and could not, be used.

## First settle which data year applies

The current MSE runs use a one-year data delay and a one-year management delay.
Under that tested timing:

1. the CMP calculation made in 2026 produces catch advice for 2027;
2. the most recent observations used are from 2025;
3. all eight CMPs use 2023–2025 data; and
4. all eight use the same five-index calculation and three-year averaging
   window.

The 2026 northern Chile acoustic estimate therefore **does not enter the 2027
calculation as it was tested in the MSE**. It would first enter the following
annual calculation, for 2028 advice.

Using observations from 2026 for the 2027 TAC would require a same-year data
rule. That would change the timing assumed in the tested CMP and should be
specified, coded, and checked before it is used. Under such a revised rule, the
three-year window would be 2024–2026.

## Data required

All eight CMPs in the named SC14 release use the same five indices.

| CMPs | Annual series required | Is the northern Chile acoustic survey used? |
|---|---|---|
| MP29, MP43, MP45, MP47, MP32, MP44, MP46, MP48 | Northern Chile acoustic survey (`Chile_AcousN`), Chile CPUE, Peru artisanal index, Peru industrial index, and offshore CPUE | Yes, when its data year falls within the approved window |

For every series, retain the value, unit, survey or fleet coverage, source,
revision status, and date received. Values must be comparable with those used
in the 2019–2023 reference period.

## Turn the observations into the CMP indicator

The raw indices are on different scales, so they cannot be averaged directly.
The current procedure performs the following calculation.

### Express each series relative to its reference-period average

For index $j$ in year $y$,

$$
R_{j,y} =
\frac{\text{index }j\text{ in year }y}
{\text{mean of index }j\text{ over 2019--2023}}.
$$

A value of 1 means that the observation equals its own 2019–2023 average. A
value of 1.25 means it is 25% above that average.

The reported acoustic estimate of about 630 kt cannot be entered directly into
the harvest rule. If its year is eligible, it must first be checked against the
same northern Chile acoustic series and converted to

$$
R_{\text{acoustic},2026} =
\frac{630\text{ kt}}
{\text{mean northern Chile acoustic biomass over 2019--2023}}.
$$

The reference-period mean is not recorded on this site, so this contribution
cannot yet be calculated here.

### Combine the available series within each year

For each year, average the relative values for the indices required by the
CMP. This gives one combined annual index $A_y$.

The analysis code averages the non-missing series. Before an operational
calculation, the Scientific Committee should decide whether a missing required
series may be omitted or whether it stops the calculation. A missing value
should never be omitted without being reported.

### Average across the recent years

The indicator supplied to the harvest rule is:

$$
I = \text{mean of the combined annual indices over the approved recent years}.
$$

For the tested 2027 timing, use 2023–2025 for all eight CMPs. One common
all-index value of $I$ is supplied to each rule.

## Apply each harvest rule

The table gives the current settings. Catch quantities use the units of the
`jmMSE` analysis; those units and the final rounding convention must be
confirmed before providing formal advice.

```{r}
#| label: tbl-annual-cmp-settings
#| tbl-cap: "Settings for the eight CMPs in the named SC14 release."
#| echo: false

annual_registry <- read.csv("../data/cmp-registry.csv",
  stringsAsFactors = FALSE, check.names = FALSE)
annual_registry <- annual_registry[
  tolower(annual_registry$sc14_release) == "yes", ]
annual_settings <- data.frame(
  CMP = paste0(annual_registry$cmp_id, " (", annual_registry$label, ")"),
  Information = "All five indices, 3 years",
  Rule = ifelse(grepl("hockeystick", annual_registry$hcr),
    "Hockey-stick", "Power ramp"),
  Minimum = annual_registry$min,
  Limit = annual_registry$lim,
  Trigger = annual_registry$trigger,
  `Target or breakpoint` = annual_registry$target,
  check.names = FALSE
)
knitr::kable(annual_settings, digits = 8)
```

In the new short-label convention, **HS** means hockeystick and **PR** means
power ramp. **+20** identifies a 20% maximum annual increase, while **-20**
identifies a 20% maximum annual decrease.

For a hockeystick CMP, the preliminary catch advice is the minimum at or below
the limit, rises in a straight line between the limit and trigger, and equals
the target at or above the trigger.

For each power-ramp CMP, let

$$
x = \frac{I-\text{limit}}{\text{trigger}-\text{limit}}.
$$

The preliminary advice is the minimum when $I$ is at or below the limit. It
is

$$
\text{minimum} + (\text{target}-\text{minimum})x^{2.5}
$$

between the limit and trigger, and

$$
\text{target}(0.2+0.8x)^{0.6}
$$

at or above the trigger. Unlike the hockeystick, the power-ramp advice can
continue to rise above its stated target.

## Apply the annual change limit

MP29 and MP32 limit the change from the previous year's catch advice to a 15%
decrease or a 20% increase. If $C^*$ is the preliminary advice and $C_{2026}$
is the agreed previous advice value, the 2027 value before rounding is

$$
C_{2027} =
\max\left(0.85C_{2026},
\min\left(C^*,1.20C_{2026}\right)\right).
$$

MP43 and MP44 instead allow a 20% decrease and a 15% increase:

$$
C_{2027} =
\max\left(0.80C_{2026},
\min\left(C^*,1.15C_{2026}\right)\right).
$$

MP45 and MP46 use symmetric 15% limits. MP47 and MP48 allow a 30% decrease and
a 20% increase. The exact multipliers are read from the CMP registry in the
worked calculation below. Under reference OM11, the realized mean
probabilities of being in the dynamic Kobe green zone over 2041–2050 were
0.6098, 0.5918, 0.6042, 0.6038, 0.6026, 0.5984, 0.5954, and 0.6056 for MP29,
MP43, MP45, MP47, MP32, MP44, MP46, and MP48, respectively.

The value 1385 appears as the initial value in the simulation controls. It
should not automatically be treated as $C_{2026}$. The operational
specification must say whether the annual limit is based on the previous TAC,
the previous CMP recommendation, or another approved quantity.

## Worked retrospective: 2025 advice using data through 2024

For this retrospective, the advice year is labelled as the year immediately
after the last data year: 2024 observations produce illustrative 2025 advice.
This simply backs the calendar labels up one year while retaining the candidate
index calculations, harvest rules, and initial comparison value of 1385.

This remains an illustration rather than advice adopted by SPRFMO. The result
is a total catch quantity in the analysis units, before an approved rounding
rule or allocation among fisheries.

```{r}
#| label: worked-calculation-setup
#| include: false

historical_indices <- read.csv(
  "../data/cmp-historical-indices.csv",
  na.strings = c("", "NA"),
  check.names = FALSE
)

all_indices <- c(
  "Chile_AcousN", "Chile_CPUE", "Peru_Artis", "Peru_Ind",
  "Offshore_CPUE"
)
reference_means <- aggregate(
  value ~ index,
  historical_indices[historical_indices$year %in% 2019:2023, ],
  function(x) mean(x, na.rm = TRUE)
)
names(reference_means)[2] <- "reference_mean"

relative_indices <- merge(
  historical_indices, reference_means, by = "index", all.x = TRUE
)
relative_indices$relative <-
  relative_indices$value / relative_indices$reference_mean

annual_relative <- function(indices) {
  selected <- relative_indices[relative_indices$index %in% indices, ]
  annual <- aggregate(
    relative ~ year, selected, function(x) mean(x, na.rm = TRUE)
  )
  annual$n_available <- vapply(annual$year, function(y) {
    sum(selected$year == y & !is.na(selected$value))
  }, integer(1))
  annual
}

recent_indicator <- function(end_year, n_years, indices) {
  annual <- annual_relative(indices)
  years <- seq(end_year - n_years + 1, end_year)
  mean(annual$relative[annual$year %in% years])
}

controls <- annual_registry
controls$label <- controls$cmp_id
controls$hcr <- sub("\\.hcr$", "", controls$hcr)
selected_mps <- c(
  "MP29", "MP43", "MP45", "MP47",
  "MP32", "MP44", "MP46", "MP48"
)
controls <- controls[match(selected_mps, controls$label), ]

calculate_advice <- function(control_row, end_year, previous = 1385) {
  indices <- all_indices
  n_years <- control_row$window_years
  indicator <- recent_indicator(end_year, n_years, indices)

  if (control_row$hcr == "hockeystick") {
    preliminary <- if (indicator <= control_row$lim) {
      control_row$min
    } else if (indicator < control_row$trigger) {
      control_row$min +
        (indicator - control_row$lim) *
        (control_row$target - control_row$min) /
        (control_row$trigger - control_row$lim)
    } else {
      control_row$target
    }
  } else if (control_row$hcr == "powerramp") {
    x <- (indicator - control_row$lim) /
      (control_row$trigger - control_row$lim)
    preliminary <- if (indicator <= control_row$lim) {
      control_row$min
    } else if (indicator < control_row$trigger) {
      control_row$min +
        (control_row$target - control_row$min) * x^2.5
    } else {
      control_row$target * (0.2 + 0.8 * x)^0.6
    }
  }

  lower <- control_row$dlow * previous
  upper <- control_row$dupp * previous
  adjusted <- max(lower, min(preliminary, upper))
  adjustment <- if (preliminary < lower) {
    paste0(round(100 * (1 - control_row$dlow)), "% decrease floor")
  } else if (preliminary > upper) {
    paste0(round(100 * (control_row$dupp - 1)), "% increase ceiling")
  } else {
    "None"
  }

  data.frame(
    CMP = control_row$label,
    `Index years` = paste0(end_year - n_years + 1, "–", end_year),
    Indicator = indicator,
    `C* before annual limit` = preliminary,
    Adjustment = adjustment,
    `Result after annual limit` = adjusted,
    check.names = FALSE
  )
}

advice_for_year <- function(end_year) {
  do.call(rbind, lapply(seq_len(nrow(controls)), function(i) {
    calculate_advice(controls[i, ], end_year = end_year)
  }))
}

advice_2024 <- advice_for_year(2024)

calculation_text <- function(control_row, indicator) {
  fmt <- function(x) formatC(x, digits = 3, format = "f")

  if (indicator <= control_row$lim) {
    return(paste0(
      "Indicator is at or below the limit; use minimum ",
      fmt(control_row$min)
    ))
  }

  if (control_row$hcr == "hockeystick") {
    if (indicator >= control_row$trigger) {
      return(paste0(
        "Indicator is at or above the trigger; use target ",
        fmt(control_row$target)
      ))
    }
    return(paste0(
      fmt(control_row$min), " + (", fmt(indicator), " − ",
      fmt(control_row$lim), ") × (", fmt(control_row$target), " − ",
      fmt(control_row$min), ") ÷ (", fmt(control_row$trigger), " − ",
      fmt(control_row$lim), ")"
    ))
  }

  x <- (indicator - control_row$lim) /
    (control_row$trigger - control_row$lim)
  if (indicator < control_row$trigger) {
    return(paste0(
      "x = (", fmt(indicator), " − ", fmt(control_row$lim), ") ÷ (",
      fmt(control_row$trigger), " − ", fmt(control_row$lim), ") = ",
      fmt(x), "; ", fmt(control_row$min), " + (",
      fmt(control_row$target), " − ", fmt(control_row$min),
      ") × x raised to 2.5"
    ))
  }

  paste0(
    "x = (", fmt(indicator), " − ", fmt(control_row$lim), ") ÷ (",
    fmt(control_row$trigger), " − ", fmt(control_row$lim), ") = ",
    fmt(x), "; ", fmt(control_row$target),
    " × (0.2 + 0.8 × x) raised to 0.6"
  )
}

rule_names <- c(
  hockeystick = "Hockeystick",
  powerramp = "Power ramp"
)

conversion_table <- do.call(rbind, lapply(seq_len(nrow(controls)), function(i) {
  data.frame(
    CMP = controls$label[i],
    Indicator = advice_2024$Indicator[i],
    Rule = unname(rule_names[controls$hcr[i]]),
    `Substitute the indicator in the rule` =
      calculation_text(controls[i, ], advice_2024$Indicator[i]),
    `C* before annual limit` = advice_2024$`C* before annual limit`[i],
    check.names = FALSE
  )
}))
```

### Historical inputs

The table contains the observations from the updated assessment data associated
with `input/0.15.dat`. The two Peruvian series are retained separately as
`Peru_Artis` and `Peru_Ind`. Each reference mean is calculated over the
available observations in 2019–2023. The northern acoustic series has no 2022
value.

```{r}
#| label: tbl-retrospective-inputs
#| tbl-cap: "Historical observations used for the retrospective calculation. A dash means that no value is stored."
#| echo: false

recent_inputs <- historical_indices[
  historical_indices$year %in% 2022:2025, ]
wide_inputs <- reshape(
  recent_inputs,
  idvar = "index",
  timevar = "year",
  direction = "wide"
)
names(wide_inputs) <- sub("^value\\.", "", names(wide_inputs))
input_table <- merge(reference_means, wide_inputs, by = "index", all.x = TRUE)
input_table <- input_table[
  match(all_indices, input_table$index),
  c("index", "reference_mean", "2022", "2023", "2024", "2025")
]
names(input_table) <- c(
  "Series", "2019–2023 mean", "2022", "2023", "2024", "2025"
)

old_na <- getOption("knitr.kable.NA")
options(knitr.kable.NA = "—")
knitr::kable(input_table, digits = 3)
options(knitr.kable.NA = old_na)
```

### Indicators for the 2025 calculation

The all-index annual values are 1.141 for 2022, 1.159 for 2023, and
1.013 for 2024. Their three-year mean is 1.104. This same indicator is supplied
to each of the eight rules.

```{r}
#| label: tbl-retrospective-indicators
#| tbl-cap: "Relative-index values supplied to the candidate harvest rules for the retrospective 2025 calculation."
#| echo: false

indicator_table <- data.frame(
  `Used by` = "All eight SC14 CMPs",
  `Index set` = "All five indices",
  `Years averaged` = "2022–2024",
  Indicator = recent_indicator(2024, 3, all_indices),
  check.names = FALSE
)

knitr::kable(indicator_table, digits = 3)
```

### Turn the indicator into preliminary advice

The indicator is a relative measure, not a catch quantity. A value of 1 means
that the recent indices are, on average, equal to their 2019–2023 reference
levels. A value above 1 means they are higher; a value below 1 means they are
lower.

Each candidate then puts its indicator into its own harvest-control rule. The
rule converts the relative indicator into a catch quantity. Because the
candidates have different limits, triggers, targets, and rule shapes, the same
indicator can produce different preliminary advice.

We write this preliminary catch advice as $C^*$. Here, $C$ means catch advice
and the star means **provisional**: it is the harvest-control-rule result before
the annual change limit is applied. The star is not another calculation.

The table shows the substitution explicitly. Values displayed in the
calculation are rounded to three decimal places; the code uses the unrounded
values.

```{r}
#| label: tbl-indicator-to-advice
#| tbl-cap: "How each candidate converts its indicator into preliminary catch advice (C*)."
#| echo: false

knitr::kable(conversion_table, digits = 3)
```

For example, MP29 uses the hockeystick rule:

$$
C^* =
270 +
\frac{(1.104-0.100)(2000-270)}
     {(2.125-0.100)}
= 1128.10.
$$

The value 1128.10 is therefore MP29's harvest-rule result. It is not yet the
final value because the annual change limit must still be applied.

### Apply the annual change limit

With 1385 as the initial comparison value, each CMP applies the decrease and
increase multipliers recorded in the registry. The table reports the range
that affects each result rather than applying one common limit to all CMPs.

```{r}
#| label: tbl-retrospective-advice
#| tbl-cap: "Illustrative 2025 total catch advice using observations through 2024 and an initial comparison value of 1385."
#| echo: false

knitr::kable(advice_2024, digits = 3)
```

These values should not be interpreted as eight equally supported
recommendations: MP43 and MP44 are the focal cases, MP29 and MP32 are
sensitivity cases, and MP45--MP48 are comparison cases. The table checks that
every rule can be reproduced from one shared set of observations and the
settings in the named registry.

## Why this example stops with 2024 observations

The updated data contain both Peruvian series through 2025, but the 2025
offshore CPUE is absent. Rather than silently omit that series or invent a
replacement, the worked calculation backs off one year and uses the complete
set available through 2024. Once the missing-data rule and 2025 offshore CPUE
status are settled, the same calculation can be advanced one year to produce
an illustrative 2026 result.

## Complete and record the recommendation

For each CMP, save:

1. the original data files and their units;
2. the 2019–2023 average for each index;
3. every annual relative index $R_{j,y}$;
4. the combined annual indices $A_y$;
5. the recent-year indicator $I$;
6. the preliminary harvest-rule result $C^*$;
7. the previous-advice value and annual-change adjustment;
8. the unrounded and rounded total recommendation;
9. any allocation calculation; and
10. all missing-data, revision, or exceptional-circumstances warnings.

The current simulations divide the total catch among fisheries using average
catch shares over the previous five years. The release check records this
implemented method consistently. It is not treated here as a further
analytical comparison; responses to review comments about allocation are
handled separately.

## What can be concluded from the currently reported value

The approximately 630 kt acoustic estimate alone is not sufficient to
calculate a TAC:

- it is outside the tested data window for a 2027 recommendation;
- it must be expressed relative to the 2019–2023 average for the same survey;
- the other four required indices are also needed.

Once the complete annual data table and the agreed previous-advice value are
available, the calculations above can be run side by side to show the
recommendation produced by every candidate.

## Implementation basis

This calculation plan is part of `SC14-MSE-2026-RC1`. It reads the eight
current settings directly from the
[CMP registry](../data/cmp-registry.csv). The [automatic agreement
report](../governance/release-check.qmd) compares that registry with the exact
500-draw simulation controls and records the analysis-code revision.
The historical index extract used in the worked calculation is saved as
[`cmp-historical-indices.csv`](../data/cmp-historical-indices.csv). It was
extracted from the updated `h1_0.16` assessment output associated with
`input/0.15.dat`, retaining both Peruvian index series.
The main implementation sources are `run_candidate_mps.R` and
`tune_candidate_change_limits.R` in the analysis revision named in that
report.
If the data delays, recent-year windows, combined-index calculation, or annual
change rule are revised, this page and the MSE tests must be updated together.
